---
title: "car_price_prediction"
output: pdf_document
date: "2024-10-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the data

```{r}
library(tidyverse)
df=read.csv("Car details v3.csv")
head(df)
dim(df)
```

Remove units from mileage and max_power
```{r}
remove_units <- function(column) {
  as.numeric(sub(" .*", "", column))
}

# Apply the function to mileage and max_power
df$mileage <- remove_units(df$mileage)
df$max_power <- remove_units(df$max_power)
df$torque <- as.character(df$torque)
df$torque <- ifelse(grepl("kgm", df$torque),
                    as.numeric(sapply(strsplit(df$torque, "[^0-9.]+"), function(x) x[1])) * 9.80665,
                    as.numeric(sapply(strsplit(df$torque, "[^0-9.]+"), function(x) x[1])))
df$engine <- as.numeric(sub(" CC", "", df$engine))
```

Remove Records with missing values
```{r}
df = na.omit(df)
dim(df)
head(df)
```
Remove outliers using IQR method

```{r}
remove_outliers <- function(df, column) {
  Q1 = quantile(df[[column]], 0.25)
  Q3 = quantile(df[[column]], 0.75)
  IQR = Q3 - Q1
  

  lower_bound = Q1 - 1.5 * IQR
  upper_bound = Q3 + 1.5 * IQR
  
  df_filtered = df %>% filter(df[[column]] >= lower_bound & df[[column]] <= upper_bound)
  
  return(df_filtered)
}

# Apply the function to remove outliers from 'selling_price'
df = remove_outliers(df, "selling_price")
dim(df)
```

Boxplot for Fuel versus selling price
```{r}
library(ggplot2)
# Assuming df is your dataframe and selling_price is in some large units
library(ggplot2)

# Convert selling_price to thousands
df$selling_price_thousands <- df$selling_price / 1000

ggplot(df, aes(x = fuel, y = selling_price_thousands, fill = fuel)) +
  geom_boxplot() +
  labs(y = "Selling Price (in thousands)") 
```

Boxplot for Fuel versus seller type
```{r}
ggplot(data = df, mapping=aes(x=seller_type, y= selling_price_thousands, fill = seller_type))+
  geom_boxplot()+
  labs(y = "Selling Price (in thousands)", x= "Seller Type") 
```

Boxplot for Fuel versus transmission type
```{r}
ggplot(data = df, mapping=aes(x=transmission, y= selling_price_thousands, fill = transmission))+
  geom_boxplot()+
  labs(y = "Selling Price (in thousands)") 
```

Boxplot for Fuel versus owner type
```{r}
ggplot(data = df, mapping=aes(x=owner, y= selling_price, fill = owner))+
  geom_boxplot()+
  labs(y = "Selling Price (in thousands)") 
```
Transform categorical features to numerical using dummy variables
```{r}
# Load the fastDummies package
library(fastDummies)
df_dummies <- dummy_cols(df, select_columns = c('fuel', 'seller_type', 'transmission', 'owner'), remove_first_dummy = TRUE)
df_dummies <- df_dummies %>% select(-fuel, -seller_type, -transmission, -owner)
head(df_dummies)
```


Correlation matrix

```{r}
df = df_dummies
numerical_columns <- df %>% select(year, selling_price, km_driven, engine, mileage, max_power, seats)

correlation_matrix <- cor(numerical_columns, use = "complete.obs")
print(correlation_matrix)
```
Scatter plot for selling_price vs max_power

```{r}

ggplot(df, aes(x = max_power, y = selling_price)) +
  geom_point(color = "blue") +
  geom_smooth(color = "red")
```

Scatter plot for engine vs max_power
```{r}
ggplot(df, aes(x = engine, y = max_power)) +
  geom_point(color = "blue") +
  geom_smooth(color = "red") +
  labs(title = "Engine Size vs Max Power", x = "Engine Size (CC)", y = "Max Power (bhp)")
```

# Scatter plot for year vs selling_price
```{r}
ggplot(df, aes(x = year, y = selling_price)) +
  geom_point(color = "blue") +
  geom_smooth(color = "red") +
  labs(title = "Year vs Selling Price", x = "Year", y = "Selling Price")
```

Scatter plot for engine vs selling_price
```{r}
ggplot(df, aes(x = engine, y = selling_price)) +
  geom_point(color = "blue") +
  geom_smooth(color = "red") +
  labs(title = "Engine Size vs Selling Price", x = "Engine Size (CC)", y = "Selling Price")
```
